\documentclass[10pt]{article}
\usepackage{336notes,fullpage,ifthen}
\parindent0em\parskip1em
\newboolean{showsols}
\pagestyle{empty}
\newenvironment{solution}{\vspace*{1em} \hrule\vspace*{0.5em} {\sf Solution.} }
                         {\vspace*{0.75em}\hrule}

\begin{document}
\setboolean{showsols}{false}
\vspace*{-5em}
\begin{center}
\large \textsf{\textbf{CAAM 336 $\cdot$ DIFFERENTIAL EQUATIONS}\\[0.5em]
Problem Set 1 \ifthenelse{\boolean{showsols}{$\cdot$ Solutions}}{}}
\end{center}

Posted Wednesday 22 August 2012.  Due Wednesday 29 August 2012, 5pm.

{\small \emph{A reminder from the course syllabus:}  
Mathematically rigorous solutions are expected; strive for clarity and elegance.
You may collaborate on the problems, but your write-up must be your own independent work.
Transcribed solutions and copied MATLAB code are both unacceptable.
\emph{You may not consult solutions from previous sections of this class.}
 Unless it is specified that a particular calculation must be performed `by hand,' you are
encouraged to use MATLAB's Symbolic Math Toolbox (or Mathematica/Wolfram Alpha/Maple)
to compute and simplify tedious integrals and derivatives on the problem sets.
As always, you must document your calculations clearly.}

\begin{enumerate}
\item {[16 points]}\\  \input classify

\vspace*{.5em}
\item {[12 points]}\\  \input solve

\vspace*{.5em}
\item {[34 points]}\\  \input checksol

\vspace*{2em}

\item {[20 points]}\\  \input steady

%\vspace*{2em}
%\item {[18 points]}\\  \input codeHat

\end{enumerate}
\end{document}
