\documentclass[11pt]{article}
\usepackage{336notes,ifthen,color}
\lfoot[\small \fancyplain{}{}]
      {\small \fancyplain{}{}}
\rfoot[\small \fancyplain{}{}]
      {\small \fancyplain{}{}}
\cfoot{\small \thepage}

\newboolean{showsols}
\newenvironment{solution}{\vspace*{1em} \hrule\vspace*{0.5em} {\sf Solution.} }
                         {\vspace*{0.75em}\hrule}

\begin{document}
\setboolean{showsols}{false}
\thispagestyle{empty}
\vspace*{0em}
   \begin{center}
      \textsf{\textbf{CAAM 336\ $\cdot$\ DIFFERENTIAL EQUATIONS}}\\[0.5em]
       \textsf{Examination~2}
   \end{center}

\vspace*{3em}

   Posted Wednesday, 5 December 2012. 

   Due no later than 5pm on Wednesday, 12 December 2012.

\vspace*{3em}
Instructions:
\begin{enumerate}
\item Time limit: \textbf{4 uninterrupted hours}.
\item There are four questions worth a total of 100 points, plus a 5 point bonus.\\
      Please do not look at the questions until you begin the exam.
\item You \emph{may not} use any outside resources,
      such as books, notes, problem sets, friends,\\
      calculators, or MATLAB. 
\item Please answer the questions thoroughly and justify all your answers.\\
      \emph{Show all your work to maximize partial credit.}

\item Print your name on the line below:

\vspace*{1em}\rule{6in}{0.5pt}

\item Time started: \hrulefill\ \ Time completed: \hrulefill

\item Indicate that this is your own individual effort in compliance with 
      the instructions above and the honor system by writing out in full 
      and signing the traditional pledge on the lines below.

\vspace*{1em}\rule{6in}{0.5pt}

\vspace*{1em}\rule{6in}{0.5pt}

\vspace*{1em}\rule{6in}{0.5pt}

\item Staple this page to the front of your exam.

\end{enumerate}

\newpage
\setcounter{page}{1}
\begin{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage\vspace*{0em}
\item {[34 points: (a),\,(c),\,(e) = 4 points; (b),\,(d) = 8 points; (f) = 6 points]}

\input fourth

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage\vspace*{0em}
\item {[22 points: (a) = 4 points; (b), (c) = 9 points]}

\input convdiff

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage\vspace*{0em}
\item {[10 points: (a) = 6 points; (b) = 4 points]}

\input couple

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage\vspace*{0em}
\item {[34 points: (a),\,(b),\,(c),\,(e),\,(f) = 4 points; (d),\,(g) = 7 points]}

\input fem2d


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage\vspace*{0em}
\item[bonus] {[5 points]}

\input bonus2

\end{enumerate}
\end{document}
